---
title: "Introduction to R and R Studio"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

# Lab 0: Sources

## Local Machine

### R

<https://cran.r-project.org/>

### R Studio

<https://posit.co/download/rstudio-desktop/>

## Posit Cloud

Section 1

<https://posit.cloud/>

## Google Colab

Section 2

<https://colab.research.google.com/notebook#create=true&language=r>

# Lab 1: Packages, Reading data and Pipe

## Packages

You may need to first install the library packages

```{r}

# install.packages("tidyverse")

# install.packages("tidyverse")

```

Libraries need to be declared every time you restart R

```{r}

library(tidyverse)
library(haven)
# library(stargazer)

```

## Reading Data

Download LFS from Odesi Nesstar first; read-in Stata datafile straight out of the archived .zip folder

### Local

```{r}

# LFS_71M0001_E_2023_May <- read_stata("LFS-71M0001-E-2023-May_F1.dta")

```

```{r}

# View(LFS_71M0001_E_2023_May)

```

### Web

```{r}

url <- "https://raw.githubusercontent.com/havaledar/ECON3740/main/LFS-71M0001-E-2023-May.zip"

# Download the ZIP file
download.file(url, destfile = '/cloud/project/LFS-71M0001-E-2023-May.zip', method = "auto")

LFS_71M0001_E_2023_May <- read_stata("LFS-71M0001-E-2023-May.zip")

```

```{r}

# View(LFS_71M0001_E_2023_May)

```

## Manipulating Data

### Pipe

```{r}

pi %>% sin %>% cos

# pi %>%
#   sin %>%
#     cos

cos(sin(pi))

```

# Lab 2

## Manipulating Data - continue

### Copy

```{r}

# Copying everything into mydata
mydata <- LFS_71M0001_E_2023_May

```

### Comment

```{r}

# This is a test

```

### Summary

```{r}

# dim(mydata)
# colnames(mydata)
# glimpse(mydata)
# head(mydata)
# summary(mydata)

```

```{r}

#Tabulate the Labour Force Status      
   
# head(mydata$LFSSTAT)
# table(mydata$LFSSTAT)
# mydata$LFSSTAT
```

```{r}

table(mydata$AGE_12)

```

```{r}

# Participation rates by gender, and by gender and year:

# head(mydata$SEX)

table(mydata$LFSSTAT, mydata$SEX)

```

```{r}

# row percentages

prop.table(table(mydata$LFSSTAT, mydata$SEX),1)

```

```{r}

# column percentages

prop.table(table(mydata$LFSSTAT, mydata$SEX),2)

```

### filter

Keep only prime age (25 to 64)

```{r}

# Investigate value labels

head(mydata$AGE_12) 

```

```{r}

# mydata %>%
#   filter(AGE_12 >= 3)

# filter(mydata, AGE_12 >= 3)

```

```{r}

d <- mydata %>%
  filter(AGE_12 >= 3 & AGE_12 <= 10)

# d <- filter(mydata, AGE_12 >= 3 & AGE_12 <= 10)

# d
```

Filter by gender

```{r}

head(d$SEX) 

```

```{r}

dM <- d %>% filter(SEX==1)
dF <- d %>% filter(SEX==2)

```

# Lab 3

## Manipulating Data - continue

### select

Keep only certain variables

```{r}

# not really needed, unless you have memory issues

d1 <- d %>% 
  select(AGE_12, SEX, LFSSTAT, EDUC, MARSTAT, PROV, CMA, MJH)

d1
```

```{r}

d1 <- d1 %>% 
  select(-CMA)

```

### Mutate

```{r}

mydata <- mydata %>% 
  mutate(NOC_10_5 = NOC_10*5)

```

```{r}

# mydata$NOC_10_5

```

### if and case

Generate the participation indicator from the labour force status variable LFSSTAT:

```{r}

head(d$LFSSTAT)

```

```{r}

# if_else

d <- d %>% 
  mutate(partic = if_else(LFSSTAT <= 3, 1, 0)) 

# d <- d %>% 
#   mutate(unemprt = case_when(LFSSTAT == 3 ~ 1,
#                              LFSSTAT == 4 ~ NA_real_,
#                              TRUE         ~ 0))


```

### Group by

Summary statistics by group

Collapsing the data by group is the easiest way of getting stats for continuous vars

```{r}
# here the group is SEX - but can use multiple groupings, eg "SEX, EDUC"

dMEAN <- d %>% 

  group_by(SEX) %>% 

  summarise(AGE_6_mean = mean(AGE_6, na.rm = T))

#view(dMEAN)

#glimpse(dMEAN)

#head(dMEAN)
```

## Basic Regression Analysis

```{r}
m1 <- lm(HRLYEARN ~ TENURE, mydata)
m1
```

```{r}
lm(log(HRLYEARN) ~ TENURE, mydata)
```

```{r}
summary(m1)
```

```{r}
m2 <- lm(partic ~ factor(EDUC),
             data=d)
summary(m2)
```
### stargazer

```{r}
install.packages("stargazer")
library(stargazer)
```

```{r}

stargazer(m2, type="text")

```
### Forecast

```{r}
#forecast
pred.yhat  <- predict(m2, d)
#out of sample forecast
pred.youts <- predict(m2, mydata)
```

## Histogram

```{r}
hist(d$HRLYEARN)
```
```{r}
hist(d$HRLYEARN[!is.na(d$HRLYEARN)])
```
```{r}
hist(d$HRLYEARN[!is.na(d$HRLYEARN)], 
     main = "Histogram of HRLYEARN (Non-NA)", 
     xlab = "HRLYEARN Values", 
     ylab = "Frequency")
```
```{r}
hist(d$HRLYEARN[!is.na(d$HRLYEARN)], 
     main = "Histogram of HRLYEARN (Non-NA)",
     xlab = "HRLYEARN Values",
     ylab = "Frequency",
     breaks = 10)
```

```{r}
hist(d$HRLYEARN[!is.na(d$HRLYEARN)], 
     main = "Histogram of HRLYEARN (Non-NA)",
     xlab = "HRLYEARN Values",
     ylab = "Frequency",
     breaks = 10,  # Specify the number of bins (in this case, 10 bins)
     col = "blue")
```
